An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullets

### 2021-09-16 CMPSpaper_writing/CMPSpaper_writing.Rmd
added caption and cleaned the comments

```{=tex}
\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\newcommand{\wj}[1]{{\textcolor{blue}{#1}}}
\setlength{\parindent}{0pt}
```
```{=tex}
\hh{Still to add: the CMPS diff measure, in which you subtract the background. 
A comparison to results from bulletxtrctr's random forest.}
\hh{The random forest is spitting out results on a [0-1] interval - so we might need to compare to the adjusted measure. That comarison should go into a separate section just before the discussion. Like - comparison to other existing methods ... special request from Alicia :)}
```
```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
library(bulletxtrctr)
library(kableExtra)
library(patchwork)
library(ggpubr)

@@

### Hamby Data Set

@@

### Extracting signatures from LEA scans

@@

### Conceptual idea of CMPS

Most algorithms for comparing striation marks are based on the digitized signatures and produce a similarity score [@song2005, @ChumbleyL_Scott2010VoTM, @aoas, @pmid30444940].
The congruent matching profile segments (CMPS) algorithm, developed by @cmps for "objective comparison of striated tool marks", is one such algorithm.
The algorithm's main idea is to take a set of consecutive and non-overlapping basis segments from the comparison and for each segment find the "best" registration position on the reference (the other bullet signature) with respect to their cross-correlation values.
From a comparison of these registration positions, a **congruent registration position** is identified, and the number of basis segments taking the congruent registration position is the CMPS score.
Note that researchers in @cmps refers the origin of basis segments as the reference, but in this paper we refer it as the comparison.
High CMPS scores are achieved between more similar signatures and are therefore indicative of a same-source pair.
Low scores between pairs of signatures are attributed to different source pairs.
However, a specific threshold of the CMPS score to distinguish between same-source and different-source comparisons is not mentioned in @cmps and depends on the underlying structure of the data and the choice of parameters.
Further research is needed to understand how to determine the threshold.
<!-- \hh{XXX What threshold did @cmps identify as the boundary between different source and same source? And what was the error rate in their example?} --> <!-- \wj{they didn't talk about a specific threshold or error rates} \hh{XXX Is there a way to infer the error based on the example? ... 'while the authors did not mention a specific threshold of the CMPS to distinguish between same-source and different-source comparison, a CMPS value of Y results in an error rate of Z in the example.'} --> The CMPS algorithm can assist firearm examiners with drawing a conclusion about the source of a comparison pair.
Unfortunately, @cmps did not release any code or specific parameter settings for their implementation of the CMPS algorithm.
Thus, in this paper we present an open-source implementation of the CMPS algorithm in the R package CMPS available from Github \url{https://github.com/willju-wangqian/CMPS} With our implementation, one can compute the CMPS score of a comparison using the following code:

@@

**Algorithm.**

@@

<!-- In the CMPS package, these three steps are implemented using a set of key functions: -->

Note that there are several parameters in the CMPS algorithm that are left to the users to decide, such as the length of basis segments `seg_length` in step 1, the number of peaks `npeaks.set` identified on each level in step 2, and the length of the tolerance zone `Tx` in both step 2 and 3.
In our implementation of the CMPS algorithm we used the parameters given in the original CMPS paper [@cmps] as the default values for these parameters.
However, the authors state that no cross-validation has been done - there might be also issues with respect to resolution of the scans.
Further research is needed, until then users are advised to think of default values as starting values and consider alternatives.
<!--Additionally, as shown in @cmps, using multi-segment lengths strategy reduces the number of "false positive" peaks in the cross-correlation curves, thus providing a more conservative way of identifying candidate positions.--> The main function that combines all steps in the CMPS algorithm described above is called `extract_feature_cmps(...)`.
Here we present it with its default parameters.

@@

In the remainder of the paper we showcase the use of the CMPS functionality on some examples and present the results of applying it to two datasets.
<!-- \hh{XXX you need a bit of an overview here on the structure of what is coming next: in the remainder of the paper we showcase the use of the CMPS functionality on some examples ...} -->

@@

<!-- ```{r, eval=FALSE} -->

<!-- install.packages("CMPS") -->

<!-- ``` -->

<!-- The development version is available from Github can be installed by using  -->

@@

<!-- \wj{XXX as mentioned before, }multi-segment lengths strategy provides a more conservative ways of finding candidate positions, so the `CMPS.score` of `cmps_with_multi_scale` is 9, while that of `cmps_without_multi_scale` is 12. \wj{XXX how should we address the advantage of multi-segment lengths strategy} \hh{XXX any score based method is walking the line between false positives and false negatives. As the number of false positives is reduced (by being more conservatives) the number of false negatives might rise. The multi scale method might also be slower because the algorithm is run a couple of times.} -->

<!-- \hh{XXX this is just a single example - a claim of 'more conservative' is quite big. At this point, it is only a claim. How would you go about finding evidence for it? } -->

@@

<!-- basis segments that agree with the congruent registration position (i.e. the segments that are congruent matching profile segments) are plotted along with the comparison signature. When plotting, each of those basis segments will be shifted to the position where the basis segment obtains a cross-correlation peak and is within the tolerance zone of the congruent registration position. -->

@@

<!-- These elements are essentially used for generating visuals above but also provide insights about the congruent registration position and the candidate positions found by each of the congruent matching profile segments.  -->

<!-- which basis segments are the congruent matching profile segments. -->

@@

```{=tex}
\begin{align}
\mathrm{CMPS_{max}} &= \max_{i,j} c_{ij} \text{ , and} \\
\mathrm{\overline{CMPS}_{max}} &= \frac{1}{n} \sum_{(i,j) \in \mathcal{P}_{k^*}} c_{ij} \text{ , where} \\
k^* &= \text{arg}\max\limits_{k} \left[  \frac{1}{n} \sum_{(i,j) \in \mathcal{P}_k} c_{ij}\right]
\end{align}
```
<!-- \begin{align} -->

<!-- \mathrm{CMPS_{max}} &= \max_{i,j} c_{ij}, \text{and }\\ -->

<!-- \mathrm{\overline{CMPS}_{max}} &= \max_{k} \left[  \frac{1}{n} \sum_{(i,j) \in \mathcal{P}_k} c_{ij}\right]. -->

<!-- \end{align} -->

<!-- \hh{XXX point out that neither of these measures evaluates the non-matches, and introduce CMPS diff here} -->

@@

\wj{Instead of computing the mean CMPS scores of matching and non-matching comparisons, a variation of $\mathrm{\overline{CMPS}_{diff}}$ is to compute the median CMPS scores of matching and non-matching comparisons. Computing the median could be beneficial if extreme values are observed for non-matching comparisons. More examples are provided in the Results section}

@@

```{=tex}
\begin{align}
c^*_{ij} = \frac{c_{ij}}{s_{ij}}
\end{align}
```
The scaled CMPS scores of the above example are shown in \autoref{fig:tiles2}(b).
Compared to the original CMPS scores, scaled scores have values within the interval $(0, 1)$ regardless of the length of the basis segments and therefore make a comparison of values possible across different parameter choices.
Similarly to the original CMPS scores, we will denote the scaled CMPS scores adjusted for out-of-phase background values by $\mathrm{\overline{CMPS^*}_{diff}}$.
For example, $\mathrm{\overline{CMPS^*}_{diff}} = 0.498$ for \autoref{fig:tiles2}(b)

@@

rf.tt$study <- c("Hamby 252", "Hamby 44")
rf.tt <- rf.tt %>% pivot_longer(
  cols = c("cmps.diff", "cmps.max", "cmps.maxbar"),
  names_to = "Metric"
)

@@

The sum of squares ratios are 20.64 and 28.96 for $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$, respectively.
\hh{This indicates, that even though scores from $\mathrm{CMPS_{max}}$ for known-match comparisons are larger than scores from the averaged version of $\mathrm{\overline{CMPS}_{max}}$, these scores achieve a better separation between the two groups of comparisons.}

@@

cols_1 <- c("#C583AE","#FFDFAA", "#748BA7", "#D9F0A0")
cols_2 <- c("#A45287","#D4AC6A", "#4C688B", "#ABC864")
# #####  Color Palette by Paletton.com
# #####  Palette URL: http://paletton.com/#uid=7561m0kllllaFw0g0qFqFg0w0aF
# 
# 
# *** Primary color:
# 
#    shade 0 = #832C65 = rgb(131, 44,101) = rgba(131, 44,101,1) = rgb0(0.514,0.173,0.396)
#    shade 1 = #C583AE = rgb(197,131,174) = rgba(197,131,174,1) = rgb0(0.773,0.514,0.682)
#    shade 2 = #A45287 = rgb(164, 82,135) = rgba(164, 82,135,1) = rgb0(0.643,0.322,0.529)
#    shade 3 = #621046 = rgb( 98, 16, 70) = rgba( 98, 16, 70,1) = rgb0(0.384,0.063,0.275)
#    shade 4 = #42002B = rgb( 66,  0, 43) = rgba( 66,  0, 43,1) = rgb0(0.259,0,0.169)
# 
# *** Secondary color (1):
# 
#    shade 0 = #AA7F39 = rgb(170,127, 57) = rgba(170,127, 57,1) = rgb0(0.667,0.498,0.224)
#    shade 1 = #FFDFAA = rgb(255,223,170) = rgba(255,223,170,1) = rgb0(1,0.875,0.667)
#    shade 2 = #D4AC6A = rgb(212,172,106) = rgba(212,172,106,1) = rgb0(0.831,0.675,0.416)
#    shade 3 = #805715 = rgb(128, 87, 21) = rgba(128, 87, 21,1) = rgb0(0.502,0.341,0.082)
#    shade 4 = #553500 = rgb( 85, 53,  0) = rgba( 85, 53,  0,1) = rgb0(0.333,0.208,0)
# 
# *** Secondary color (2):
# 
#    shade 0 = #2B4970 = rgb( 43, 73,112) = rgba( 43, 73,112,1) = rgb0(0.169,0.286,0.439)
#    shade 1 = #748BA7 = rgb(116,139,167) = rgba(116,139,167,1) = rgb0(0.455,0.545,0.655)
#    shade 2 = #4C688B = rgb( 76,104,139) = rgba( 76,104,139,1) = rgb0(0.298,0.408,0.545)
#    shade 3 = #143054 = rgb( 20, 48, 84) = rgba( 20, 48, 84,1) = rgb0(0.078,0.188,0.329)
#    shade 4 = #051B38 = rgb(  5, 27, 56) = rgba(  5, 27, 56,1) = rgb0(0.02,0.106,0.22)
# 
# *** Complement color:
# 
#    shade 0 = #81A035 = rgb(129,160, 53) = rgba(129,160, 53,1) = rgb0(0.506,0.627,0.208)
#    shade 1 = #D9F0A0 = rgb(217,240,160) = rgba(217,240,160,1) = rgb0(0.851,0.941,0.627)
#    shade 2 = #ABC864 = rgb(171,200,100) = rgba(171,200,100,1) = rgb0(0.671,0.784,0.392)
#    shade 3 = #5B7814 = rgb( 91,120, 20) = rgba( 91,120, 20,1) = rgb0(0.357,0.471,0.078)
#    shade 4 = #395000 = rgb( 57, 80,  0) = rgba( 57, 80,  0,1) = rgb0(0.224,0.314,0)
# 
# 
# #####  Generated by Paletton.com (c) 2002-2014
```

@@

labels <- expression(bar(CMPS)[diff], {bar(CMPS)^symbol("*")}[diff], CMPS[max], bar(CMPS)[max])

@@

p.npeak_plot <- table.cmps_neapk.long %>%
  ggplot(aes(x = xstart + 3*as.numeric(factor(Metric))-9)) +
  geom_segment(aes(xend = xstart + 3*as.numeric(factor(Metric))-9, 
                   y = value, yend= 0, color=Metric),
               position=position_dodge(width = 0.6)) +
  geom_point(aes(y =  value, color = Metric), 
             alpha = 0.8,
             position=position_dodge(width = 0.5), size = 3.5) +
  facet_grid(.~study) +
  scale_x_continuous(breaks = npeak_convert$xstart, labels = npeak_convert$npeak ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "bottom") +
  ylab("Sums of Squares Ratio") + 
  xlab("Number of Peaks") +
  scale_colour_manual(values=cols_2, labels = labels) +
  scale_fill_manual(values=cols_1, labels = labels) 
```

@@

```{r result1_44, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; outliers are removed in bullet signatures; \\texttt{seg\\_length = 61}, \\texttt{Tx = 30}, \\texttt{npeaks.set = c(5,3,1)} ", fig.height=3}
# [4] "span1=0.25, span2=0.03, rm_outliers = TRUE, 5025, 531"  
# knitr::include_graphics("img/hamby252_v1.png", dpi = 100)
# CMPS_hamby252_results$plot[[2]]

p1 <- metric_plot_helper(CMPS_hamby44_results_seg$cmps.table[[3]], "cmps.max",
                         subtitle = "npeaks.set: 5-3-1, seg_length: 61") +
  xlab(labels[3])
p2 <- metric_plot_helper(CMPS_hamby44_results_seg$cmps.table[[3]], "cmps.maxbar",
                         subtitle = "npeaks.set: 5-3-1, seg_length: 61") +
  xlab(labels[4])
p.result1_44 <- ggarrange(plotlist = list(p1, p2), nrow = 1, ncol = 2,
               common.legend = TRUE, legend = "bottom")
p.result1_44 <- annotate_figure(p.result1_44, top = text_grob(com.title44))
p.result1_44
```

@@

```{r param_seg_plot, echo=FALSE, out.width="400px", fig.cap="Comparison of results from the CMPS algorithm based on different basis segment lengths (parameter \\texttt{seg\\_length}). Only the $\\mathrm{CMPS_{max}}$ metric suggests that the default values for the basis segment length result in the best separation. Better separation is achieved based on the modified CMPS metrics, inclusing the newly suggested ones. For Hamby 252 these metrics agree that a segment length of 75 and a segment length of 122 for Hamby 44 yield better results", fig.height=3, warning = FALSE}
p.seg_plot 
```

```{r param_npeak_plot, echo=FALSE, out.width="400px", fig.cap="Comparison of CMPS results based on different strategies of number of peak selections. \\hh{5-3-1 should be next to 5-3-1*}. Starred results compare CMPS performance with results published in the literature. ", fig.height=3, warning = FALSE}
p.npeak_plot
```

@@

@cmps present histograms of $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$ for `npeaks.set = 5` and `npeaks.set = c(5,3,1)` with (presumably) Hamby 252.
The values in these histograms allow us to calculate the sum of squares ratios and include the results in \autoref{fig:param_npeak_plot} as well.
They are marked by an asterisk at the top right corner in \autoref{fig:param_npeak_plot}.
The sum of squares ratios we obtained for `npeaks.set = 5` and `npeaks.set = c(5,3,1)` is slightly higher than those obtained from the histograms of @cmps. 
\wj{One interesting observation is that cmps.max works pretty well for 531 in nist's paper... do we need to include this?XXX} \hh{no need}
Since the researchers of @cmps did not use $\mathrm{\overline{CMPS}_{diff}}$ or $\mathrm{\overline{CMPS^*}_{diff}}$, and they did not apply the CMPS algorithm to the Hamby set 44, we were not able to compare those results.

In \autoref{fig:param_npeak_plot}, we also included the sum of squares ratios computed from the random forest scores [@aoas] for different metrics. 
<!--A random forest model can be used for both classification and regression. 
It is an ensemble learning method that takes advantage of multiple predictive features.-->
The random forest model presented in @aoas was trained at the Center for Statistics and Applications in Forensic Evidence (CSAFE) and is publicly available in the R package `bulletxtrctr` [@bulletxtrctr].
Similarly to the CMPS algorithm, this random forest model produces a  score to quantify the similarity of a land-by-land comparison.
The RF scores lie in an interval of [0,1] making them most comparable to the scaled CMPS scores.
We applied the logic of $\mathrm{{CMPS}_{max}}$, $\mathrm{\overline{CMPS}_{max}}$, and $\mathrm{\overline{CMPS}_{diff}}$ to the random forest scores and obtained results of the random forest model for different metrics. 
As shown in \autoref{fig:param_npeak_plot}, the random forest model does not achieve the sum of squares ratio as high as those achieved by the top CMPS algorithms for Hamby 252, but it does achieve better results for Hamby 44.
This might suggest that an inclusion of the CMPS score as an additional feature in the random forest model training might be beneficial for an overall separation to determine source.

@@

\hh{XXX Which 'It' are you referring to?}
It would also facilitate future studies on parameter cross-validations, method comparisons, and the development of statistical methods for modeling KM and KNM CMPS score distributions, which can potentially be used to estimate error rates or define an identification criterion.

>>>>

```{=tex}
\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\newcommand{\wj}[1]{{\textcolor{blue}{#1}}}
\setlength{\parindent}{0pt}
```

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
library(bulletxtrctr)
library(kableExtra)
library(patchwork)
library(ggpubr)

@@

**Hamby Data Set**
<!-- ### Hamby Data Set -->

@@

**Extracting signatures from LEA scans**
<!-- ### Extracting signatures from LEA scans -->

@@

**Conceptual idea of CMPS**
<!-- ### Conceptual idea of CMPS -->

Most algorithms for comparing striation marks are based on the digitized signatures and produce a similarity score [@song2005, @ChumbleyL_Scott2010VoTM, @aoas, @pmid30444940].
The congruent matching profile segments (CMPS) algorithm, developed by @cmps for "objective comparison of striated tool marks", is one such algorithm.
The algorithm's main idea is to take a set of consecutive and non-overlapping basis segments from the comparison and for each segment find the "best" registration position on the reference (the other bullet signature) with respect to their cross-correlation values.
From a comparison of these registration positions, a **congruent registration position** is identified, and the number of basis segments taking the congruent registration position is the CMPS score.
Note that researchers in @cmps refers the origin of basis segments as the reference, but in this paper we refer it as the comparison.
High CMPS scores are achieved between more similar signatures and are therefore indicative of a same-source pair.
Low scores between pairs of signatures are attributed to different source pairs.
However, a specific threshold of the CMPS score to distinguish between same-source and different-source comparisons is not mentioned in @cmps and depends on the underlying structure of the data and the choice of parameters.
Further research is needed to understand how to determine the threshold.
The CMPS algorithm can assist firearm examiners with drawing a conclusion about the source of a comparison pair.
Unfortunately, @cmps did not release any code or specific parameter settings for their implementation of the CMPS algorithm.
Thus, in this paper we present an open-source implementation of the CMPS algorithm in the R package CMPS available from Github \url{https://github.com/willju-wangqian/CMPS} With our implementation, one can compute the CMPS score of a comparison using the following code:

@@

**Algorithm**

@@

Note that there are several parameters in the CMPS algorithm that are left to the users to decide, such as the length of basis segments `seg_length` in step 1, the number of peaks `npeaks.set` identified on each level in step 2, and the length of the tolerance zone `Tx` in both step 2 and 3.
In our implementation of the CMPS algorithm we used the parameters given in the original CMPS paper [@cmps] as the default values for these parameters.
However, the authors state that no cross-validation has been done - there might be also issues with respect to resolution of the scans.
Further research is needed, until then users are advised to think of default values as starting values and consider alternatives.
The main function that combines all steps in the CMPS algorithm described above is called `extract_feature_cmps(...)`.
Here we present it with its default parameters.

@@

In the remainder of the paper we showcase the use of the CMPS functionality on some examples and present the results of applying it to two datasets.

@@

@@

@@

@@

@@

```{=tex}
\begin{align}
\mathrm{CMPS_{max}} &= \max_{i,j} c_{ij} \text{ , and} \\
\mathrm{\overline{CMPS}_{max}} &= \frac{1}{n} \sum_{(i,j) \in \mathcal{P}_{k^*}} c_{ij} \text{ , where} \\
k^* &= \text{arg}\max\limits_{k} \left[  \frac{1}{n} \sum_{(i,j) \in \mathcal{P}_k} c_{ij}\right]
\end{align}
```

@@

<!-- ??? We shouldn't talk about the median version, should we? -->

<!-- \wj{Instead of computing the mean CMPS scores of matching and non-matching comparisons, a variation of $\mathrm{\overline{CMPS}_{diff}}$ is to compute the median CMPS scores of matching and non-matching comparisons. Computing the median could be beneficial if extreme values are observed for non-matching comparisons. More examples are provided in the Results section} -->

@@

```{=tex}
\begin{align}
c^*_{ij} = \frac{c_{ij}}{s_{ij}}
\end{align}
```
The scaled CMPS scores of the above example are shown in \autoref{fig:tiles2}(b).
Compared to the original CMPS scores, scaled scores have values within the interval $[0, 1]$ regardless of the length of the basis segments and therefore make a comparison of values possible across different parameter choices.
Similarly to the original CMPS scores, we will denote the scaled CMPS scores adjusted for out-of-phase background values by $\mathrm{\overline{CMPS^*}_{diff}}$.
For example, $\mathrm{\overline{CMPS^*}_{diff}} = 0.498$ for \autoref{fig:tiles2}(b)

@@

rf.tt$cmps.diff_scaled <- rf.tt$cmps.diff

rf.tt$study <- c("Hamby 252", "Hamby 44")
rf.tt <- rf.tt %>% pivot_longer(
  cols = c("cmps.diff", "cmps.max", "cmps.maxbar", "cmps.diff_scaled"),
  names_to = "Metric"
)

@@

The sum of squares ratios are 20.64 and 28.96 for $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$, respectively.
This indicates, that even though scores from $\mathrm{CMPS_{max}}$ for known-match comparisons are larger than scores from the averaged version of $\mathrm{\overline{CMPS}_{max}}$, these scores achieve a better separation between the two groups of comparisons.

@@

cols_1 <- c("#748BA7", "#C583AE","#FFDFAA", "#D9F0A0")
cols_2 <- c("#4C688B", "#A45287","#D4AC6A", "#ABC864")
# #####  Color Palette by Paletton.com
# #####  Palette URL: http://paletton.com/#uid=7561m0kllllaFw0g0qFqFg0w0aF
# 
# 
# *** Primary color:
# 
#    shade 0 = #832C65 = rgb(131, 44,101) = rgba(131, 44,101,1) = rgb0(0.514,0.173,0.396)
#    shade 1 = #C583AE = rgb(197,131,174) = rgba(197,131,174,1) = rgb0(0.773,0.514,0.682)
#    shade 2 = #A45287 = rgb(164, 82,135) = rgba(164, 82,135,1) = rgb0(0.643,0.322,0.529)
#    shade 3 = #621046 = rgb( 98, 16, 70) = rgba( 98, 16, 70,1) = rgb0(0.384,0.063,0.275)
#    shade 4 = #42002B = rgb( 66,  0, 43) = rgba( 66,  0, 43,1) = rgb0(0.259,0,0.169)
# 
# *** Secondary color (1):
# 
#    shade 0 = #AA7F39 = rgb(170,127, 57) = rgba(170,127, 57,1) = rgb0(0.667,0.498,0.224)
#    shade 1 = #FFDFAA = rgb(255,223,170) = rgba(255,223,170,1) = rgb0(1,0.875,0.667)
#    shade 2 = #D4AC6A = rgb(212,172,106) = rgba(212,172,106,1) = rgb0(0.831,0.675,0.416)
#    shade 3 = #805715 = rgb(128, 87, 21) = rgba(128, 87, 21,1) = rgb0(0.502,0.341,0.082)
#    shade 4 = #553500 = rgb( 85, 53,  0) = rgba( 85, 53,  0,1) = rgb0(0.333,0.208,0)
# 
# *** Secondary color (2):
# 
#    shade 0 = #2B4970 = rgb( 43, 73,112) = rgba( 43, 73,112,1) = rgb0(0.169,0.286,0.439)
#    shade 1 = #748BA7 = rgb(116,139,167) = rgba(116,139,167,1) = rgb0(0.455,0.545,0.655)
#    shade 2 = #4C688B = rgb( 76,104,139) = rgba( 76,104,139,1) = rgb0(0.298,0.408,0.545)
#    shade 3 = #143054 = rgb( 20, 48, 84) = rgba( 20, 48, 84,1) = rgb0(0.078,0.188,0.329)
#    shade 4 = #051B38 = rgb(  5, 27, 56) = rgba(  5, 27, 56,1) = rgb0(0.02,0.106,0.22)
# 
# *** Complement color:
# 
#    shade 0 = #81A035 = rgb(129,160, 53) = rgba(129,160, 53,1) = rgb0(0.506,0.627,0.208)
#    shade 1 = #D9F0A0 = rgb(217,240,160) = rgba(217,240,160,1) = rgb0(0.851,0.941,0.627)
#    shade 2 = #ABC864 = rgb(171,200,100) = rgba(171,200,100,1) = rgb0(0.671,0.784,0.392)
#    shade 3 = #5B7814 = rgb( 91,120, 20) = rgba( 91,120, 20,1) = rgb0(0.357,0.471,0.078)
#    shade 4 = #395000 = rgb( 57, 80,  0) = rgba( 57, 80,  0,1) = rgb0(0.224,0.314,0)
# 
# 
# #####  Generated by Paletton.com (c) 2002-2014
```

@@

labels <- expression( CMPS[max], bar(CMPS)[diff], {bar(CMPS)^symbol("*")}[diff], bar(CMPS)[max])

@@

table.cmps_neapk.long <- table.cmps_neapk.long %>% mutate(
  rf_status = if_else(npeak == "RF score* (Hare et al)",
                      TRUE, FALSE)
)

table.cmps_neapk.long$Metric <- 
  table.cmps_neapk.long$Metric %>% 
  factor(levels = c("cmps.max", "cmps.diff", "cmps.diff_scaled", "cmps.maxbar"))

p.npeak_plot <- table.cmps_neapk.long %>%
  ggplot(aes(x = xstart + 3*as.numeric(factor(Metric))-9)) +
  geom_segment(aes(xend = xstart + 3*as.numeric(factor(Metric))-9, 
                   y = value, yend= 0, color=Metric),
               position=position_dodge(width = 0.6)) +
  geom_point(aes(y =  value, color = Metric, shape = rf_status), 
             alpha = 0.8,
             position=position_dodge(width = 0.5), size = 3.5) +
  facet_grid(.~study) +
  ylab("Sums of Squares Ratio") + 
  xlab("Number of Peaks") +
  coord_flip() +
  scale_x_continuous(breaks = npeak_convert$xstart, labels = npeak_convert$npeak ) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(shape = FALSE) + 
  scale_shape_manual(values = c(16, 1)) +
  scale_colour_manual(values=cols_2, labels = labels) +
  scale_fill_manual(values=cols_1, labels = labels) 
```

@@

```{r result1_44, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; outliers are removed in bullet signatures; \\texttt{seg\\_length = 61}, \\texttt{Tx = 30}, \\texttt{npeaks.set = c(5,3,1)} ", fig.height=3}
p1 <- metric_plot_helper(CMPS_hamby44_results_seg$cmps.table[[3]], "cmps.max",
                         subtitle = "npeaks.set: 5-3-1, seg_length: 61") +
  xlab(labels[3])
p2 <- metric_plot_helper(CMPS_hamby44_results_seg$cmps.table[[3]], "cmps.maxbar",
                         subtitle = "npeaks.set: 5-3-1, seg_length: 61") +
  xlab(labels[4])
p.result1_44 <- ggarrange(plotlist = list(p1, p2), nrow = 1, ncol = 2,
               common.legend = TRUE, legend = "bottom")
p.result1_44 <- annotate_figure(p.result1_44, top = text_grob(com.title44))
p.result1_44
```

@@

```{r param_seg_plot, echo=FALSE, out.width="400px", fig.cap="Comparison of results from the CMPS algorithm based on different basis segment lengths (parameter \\texttt{seg\\_length}). Only the $\\mathrm{CMPS_{max}}$ metric suggests that the default values for the basis segment length result in the best separation. Better separation is achieved based on the modified CMPS metrics, inclusing the newly suggested ones. For Hamby 252 these metrics agree on a segment length of 75, and a segment length of 122 for Hamby 44 yields better results", fig.height=3, warning = FALSE}
p.seg_plot 
```

```{r param_npeak_plot, echo=FALSE, out.width="400px", fig.cap="Comparison of CMPS results based on different strategies of number of peak selections. Starred results compare CMPS performance with results published in the literature. Results for the random forest score are represented with circles because the metrics are computed not based on the CMPS scores, but the random forest scores with the same logic. Since random forest scores lie within the interval $[0, 1]$, scaling the random forest scores will not change the results.", fig.height=3, warning = FALSE}
p.npeak_plot
```

@@

@cmps present histograms of $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$ for `npeaks.set = 5` and `npeaks.set = c(5,3,1)` with (presumably) Hamby 252.
The values in these histograms allow us to calculate the sum of squares ratios and include the results in \autoref{fig:param_npeak_plot} as well.
They are marked by an asterisk at the top right corner in \autoref{fig:param_npeak_plot}.
The sum of squares ratios we obtained for `npeaks.set = 5` and `npeaks.set = c(5,3,1)` is slightly higher than those obtained from the histograms of @cmps. 
It's curious to see that for the Hamby 252 results published in @cmps the $\mathrm{{CMPS}_{max}}$ metric achieves values of the sum of squares ratio higher than those achieved by the $\mathrm{\overline{CMPS}_{max}}$ metric.

Since the researchers of @cmps did not use $\mathrm{\overline{CMPS}_{diff}}$ or $\mathrm{\overline{CMPS^*}_{diff}}$, and they did not apply the CMPS algorithm to the Hamby set 44, we were not able to compare those results.

In \autoref{fig:param_npeak_plot}, we also included the sum of squares ratios computed from the random forest scores [@aoas] for different metrics. 
The random forest model presented in @aoas was trained at the Center for Statistics and Applications in Forensic Evidence (CSAFE) and is publicly available in the R package `bulletxtrctr` [@bulletxtrctr].
Similarly to the CMPS algorithm, this random forest model produces a  score to quantify the similarity of a land-by-land comparison.
The RF scores lie in an interval of [0,1] making them most comparable to the scaled CMPS scores.
We applied the logic of $\mathrm{{CMPS}_{max}}$, $\mathrm{\overline{CMPS}_{max}}$, and $\mathrm{\overline{CMPS}_{diff}}$ to the random forest scores and obtained results of the random forest model for different metrics. 
As shown in \autoref{fig:param_npeak_plot}, the random forest model does not achieve the sum of squares ratio as high as those achieved by the top CMPS algorithms for Hamby 252, but it does achieve better results for Hamby 44.
This might suggest that an inclusion of the CMPS score as an additional feature in the random forest model training might be beneficial for an overall separation to determine source.

@@

Open-source implementations not only allow a peer-review but also facilitate further research, such as parameter cross-validations, method comparisons, and the development of statistical methods for modeling KM and KNM CMPS score distributions, which can then be used for error-rate estimations and fair applications that were called for by the PCAST [@pcast] report.
